### 2020-08-11 latex/vnhcs.tex
Wrote the proof of regularity, but it's not compiling so that needs fixing.

% TODO: Replace [p D_i p] here with our new notation
The resulting Hamiltonian system reduces to (\ref{eqn:hamiltonian}).
\begin{align}\label{eqn:hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2} p\tpose D^{-1}(q) p + P(q) \\
                     &\begin{cases}
        \dot{q} = D\inv(q)p \\
        \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q D\inv(q) p
        - \nabla_q P(q) + B(q) \tau \\
    \end{cases} 
\end{align}

@@

% TODO: Replace the [P M_i P] with new notation
\begin{thm}
    Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
    \ref{assm:B-orthogonal}, the Hamiltonian system (\ref{eqn:hamiltonian}) is
    simply actuated with simply actuated coordinates 
    \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\). The resulting dynamics are 
    given by (\ref{eqn:simple-hamiltonian}),
    \begin{align}\label{eqn:simple-hamiltonian}
        \hat{H}(Q,P) &= 
        \frac{1}{2} P\tpose \Minv(Q) P + V(Q) \\
       &\begin{cases}
            \dot{Q} = \Minv(Q)P \\
            \dot{P} = -\frac{1}{2} (I_n \otimes P\tpose) \nabla_Q \Minv(Q) P
                - \nabla_Q V(Q) + \simpleB \tau
        \end{cases} \nonumber
    \end{align}
    where
    \begin{align*}
        \Minv(Q) &:= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose \\
        V(Q) &:= P(\mathbf{B}\tpose Q) \\
    \end{align*}
\end{thm}
\begin{proof}
    The Poisson bracket between the functions \(f(q,p)\) and \(g(q,p)\) is defined by
    \cite{landau_mechanics} as follows:
    \[
        [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
                \pdiff{f}{q_i}\pdiff{g}{p_i}
    \]
    For any constant matrix \(A\), the transformation
    \(\{Q = Aq, P = Ap\}\) satisfies
    \(\pdiff{Q_i}{p_m} = \pdiff{P_i}{q_m} = 0\) for all 
    \(i,m \in \n\).
    Hence, for our new coordinates \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\),
    \begin{align*}
        [Q_i,Q_j] &:= \sum\limits_{m = 1}^n \pdiff{Q_i}{p_m}\pdiff{Q_j}{q_m} - 
        \pdiff{Q_i}{q_m}\pdiff{Q_j}{p_m} = 0 \\
        [P_i,P_j] &:= \sum\limits_{m=1}^n \pdiff{P_i}{p_m}\pdiff{P_j}{q_m} -
        \pdiff{P_i}{q_m}\pdiff{P_j}{p_m} = 0
    \end{align*}
    Since the matrix \(\mathbf{B}\) is orthogonal,
    \((\mathbf{B}_i)\tpose \mathbf{B}\tpose_j = (\mathbf{B}_i)\tpose (\mathbf{B}^{-1})_j = \delta_{i,j}\). Using this
    fact we see that the Poisson brackets between \(P\) and \(Q\) are given by:
    \begin{align*}
        [P_i,Q_j] &= \sum\limits_{m=1}^n\pdiff{P_i}{p_m}\pdiff{Q_j}{q_m}
        - \pdiff{P_i}{q_m}\pdiff{Q_j}{p_m} \\
                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}_{j,m} - 0 \\
                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}\tpose_{m,j} \\
                  &= (\mathbf{B}_i)\tpose \mathbf{B}\tpose_j \\
                  &= \delta_{i,j}
    \end{align*}

@@

% TODO: Write the [p' M p] matrix in our nice nottation
\begin{assm}\label{assm:H-is-simply-actuated}
    The mechanical system under consideration is a
    Hamiltonian system with \(n\) degrees of freedom and 
    \(k < n\) actuators. It is described in simply
    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\) and has the
    following dynamics:
    \begin{align*}
        \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q) \\
         &\begin{cases}
            \dot{q} = \Minv p \\
            \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q \Minv(q) p
            - \nabla_q V(q) + \simpleB \tau \\
        \end{cases}
    \end{align*}
\end{assm}

\begin{thm}\label{thm:vnhc-regularity}
    A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
    and
    \[
        \rank{\left(dh_q \Minv(q) + 
          dh_{p_u} (I_n \otimes p\tpose)\nabla_{q_u}\Minv(q) 
         \right)\simpleB} = k
    \]
    everywhere on the constraint manifold \(\Gamma\).
\end{thm}
\begin{proof}
    % TODO: complete this proof with our notation
\end{proof}

From the proof for Theorem \ref{thm:vnhc-regularity}, a regular VNHC of order
\(k\) can be stabilized by the output-linearizing phase-feedback controller
(\ref{eqn:vnhc-torque-controller}).
% TODO: Controller for VNHC in closed form
\begin{equation}\label{eqn:vnhc-torque-controller}
    \tau(q,p) = ? - k_p e - k_d \dot{e}
\end{equation}
where \(k_p, k_d \in \R_{>0}\) can be tuned using standard linear control
techniques.

@@

Now that we have found a controller to enforce a regular VNHC of order \(k\), we
would like to solve for the closed-loop dynamics. Naturally, these dynamics
should be specified through \((q_u, p_u)\) since \(q_a\) is a function of these
as specified by \(h(q,p_u) = 0\).
Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). This is because
the \(\dot{p}\) dynamics contains the coupling term 
\((I_n \otimes p\tpose)\nabla_{q_u}M(q)p\).

@@

\begin{thm}\label{thm:zero-dynamics}
    Let \(\mathcal{H}\) be a ----- mechanical system satisfying Assumption
    \ref{assm:H-is-simply-actuated}. Let
    \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint manifold
    \(\Gamma\).
    Then, on \(\Gamma\), the closed-loop dynamics are given by
    \begin{equation}\label{eqn:qpu-dynamics}
        \left.\begin{aligned}
            \dot{q}_u &= M(q)p \\
            \dot{p}_u &= -\nabla_{q_u}V(q) \\
            \end{aligned}\right|_{\begin{array}{c}
                h(q,p_u) = 0 \\ 
                pa = g(q_u,p_u) \\
            \end{array}}
    \end{equation}
    where
    % TODO: This is too wide, shrink it somehow
    \begin{equation}\label{eqn:g-qpu}
        g(q_u,p_u) := \left.\left(dh_q \Minv(q) \simpleB \right)\inv 
        \left(dh_{p_u} \nabla_{q_u}V(q) - dh_q \Minv(q)
        \begin{bmatrix}
            I_{n-k} \\
            \Zmat{k \times (n-k)} \\
        \end{bmatrix} p_u\right)\right|_{h(q,p_u) = 0}
    \end{equation}
\end{thm}
\begin{proof}
%TODO: Complete the proof for this theorem
\end{proof}

@@

>>>>

% TODO: Replace [p D_i p] here with our new notation
The resulting Hamiltonian system reduces to (\ref{eqn:hamiltonian}).
\begin{align}\label{eqn:hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2} p\tpose D^{-1}(q) p + P(q) \\
                     &\begin{cases}
        \dot{q} = D\inv(q)p \\
        \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
        - \nabla_q P(q) + B(q) \tau \\
    \end{cases} 
\end{align}

@@

% TODO: Replace the [P M_i P] with new notation
\begin{thm}
    Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
    \ref{assm:B-orthogonal}, the Hamiltonian system (\ref{eqn:hamiltonian}) is
    simply actuated with simply actuated coordinates 
    \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\). The resulting dynamics are 
    given by (\ref{eqn:simple-hamiltonian}),
    \begin{align}\label{eqn:simple-hamiltonian}
        \hat{H}(Q,P) &= 
        \frac{1}{2} P\tpose \Minv(Q) P + V(Q) \\
       &\begin{cases}
            \dot{Q} = \Minv(Q)P \\
            \dot{P} = -\frac{1}{2} (\Id{n} \otimes P\tpose) \nabla_Q \Minv(Q) P
                - \nabla_Q V(Q) + \simpleB \tau
        \end{cases} \nonumber
    \end{align}
    where
    \begin{align*}
        \Minv(Q) &:= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose \\
        V(Q) &:= P(\mathbf{B}\tpose Q) \\
    \end{align*}
\end{thm}
\begin{proof}
    The Poisson bracket between the functions \(f(q,p)\) and \(g(q,p)\) is defined by
    \cite{landau_mechanics} as follows:
    \[
        [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
                \pdiff{f}{q_i}\pdiff{g}{p_i}
    \]
    For any constant matrix \(A\), the transformation
    \(\{Q = Aq, P = Ap\}\) satisfies
    \(\pdiff{Q_i}{p_m} = \pdiff{P_i}{q_m} = 0\) for all 
    \(i,m \in \n\).
    Hence, for our new coordinates \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\),
    \begin{align*}
        [Q_i,Q_j] &:= \sum\limits_{m = 1}^n \pdiff{Q_i}{p_m}\pdiff{Q_j}{q_m} - 
        \pdiff{Q_i}{q_m}\pdiff{Q_j}{p_m} = 0 \\
        [P_i,P_j] &:= \sum\limits_{m=1}^n \pdiff{P_i}{p_m}\pdiff{P_j}{q_m} -
        \pdiff{P_i}{q_m}\pdiff{P_j}{p_m} = 0
    \end{align*}
    Since the matrix \(\mathbf{B}\) is orthogonal,
    \((\mathbf{B}_i)\tpose \mathbf{B}\tpose_j = (\mathbf{B}_i)\tpose (\mathbf{B}^{-1})_j = \delta_{i,j}\). Using this
    fact we see that the Poisson brackets between \(P\) and \(Q\) are given by:
    \begin{align*}
        [P_i,Q_j] &= \sum\limits_{m=1}^n\pdiff{P_i}{p_m}\pdiff{Q_j}{q_m}
        - \pdiff{P_i}{q_m}\pdiff{Q_j}{p_m} \\
                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}_{j,m} - 0 \\
                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}\tpose_{m,j} \\
                  &= (\mathbf{B}_i)\tpose \mathbf{B}\tpose_j \\
                  &= \delta_{i,j}
    \end{align*}

@@

\begin{assm}\label{assm:H-is-simply-actuated}
    The mechanical system under consideration is a
    Hamiltonian system with \(n\) degrees of freedom and 
    \(k < n\) actuators. It is described in simply
    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\) and has the
    following dynamics:
    \begin{align*}
        \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q) \\
         &\begin{cases}
            \dot{q} = \Minv p \\
            \dot{p} = -\frac{1}{2} \pdmat - \nabla_q V(q) + \simpleB \tau \\
        \end{cases}
    \end{align*}
\end{assm}
\begin{notation}
    We will write \(q_u \in \mathcal{Q}_u\), \(q_a \in \mathcal{Q}_a\) where
    \(\mathcal{Q}_u \times \mathcal{Q}_a = \mathcal{Q}\). 
    We also write
    \(p_u \in \mathcal{P}_u = \R^{n-k}\) and 
    \(p_a \in \mathcal{P}_a = \R^k\), so that 
    \(p \in \mathcal{P} := \mathcal{P}_u \times \mathcal{P}_a\). 
    In this manner, the phase space of our system can be written as
    \(\mathcal{Q} \times \mathcal{P}\).
\end{notation}

\begin{thm}\label{thm:vnhc-regularity}
    A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
    and
    \[
        \rank{\left(dh_q \Minv(q) - 
          dh_{p_u} (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
         \right)\simpleB} = k
    \]
    everywhere on the constraint manifold \(\Gamma\).
\end{thm}
\begin{proof}
    % TODO: complete this proof with our notation
    Let \(e = h(q,p) \in \R^k\). Then 
    \begin{align*}
        \dot{e} &= dh_q \dot{q} + dh_p \dot{p} \\
                &= dh_q \Minv(q)p +{}  \\
            & \begin{bmatrix} dh_{p_u} & dh_{p_a} \end{bmatrix}
        \left( -\frac{1}{2} \begin{bmatrix}
            (\Id{n-k} \otimes p\tpose) \nabla_{q_u}\Minv(q)p \\
            (\Id{k} \otimes p\tpose) \nabla_{q_a}\Minv(q)p
            \end{bmatrix} - \begin{bmatrix}
            \nabla_{q_u}V(q) \\
            \nabla_{q_a}V(q)
        \end{bmatrix} + \simpleB \tau\right)
    \end{align*}
    If \(dh_{p_a} \neq \Zmat{k \times k}\) for some \((q,p)\) on \(\Gamma\), 
    then \(\tau\) appears in \(\dot{e}\) and the VNHC is not of relative degree
    \(\{2,2,\ldots,2\}\).
    Hence, we must have that \(dh_{p_a} = \Zmat{k \times k}\) if we want the
    VNHC to be regular. Proceeding with this assumption, we now find that
    \(h : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R^k\), which means
    \[
        \dot{e} = dh_q \Minv(q)p - 
        dh_{p_u} \left(\frac{1}{2} \pudmat + \nabla_{q_u}V(q)\right)
    \]
    Taking one further derivative provides
    \begin{align*}
        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q) p + 
        dh_q \left(\sum\limits_{i=1}^n \pdiff{\Minv}{q_i}(q)\dot{q_i}\right)p + 
        dh_q \Minv(q) \dot{p} - \\
         & \diff{}{t}\left\{dh_{p_u}\right\}
         \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) - \\
         & dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left\{\pudmat\right\} + 
         \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} \right)
    \end{align*}
    We will compute the explicit expression of \(\ddot{e}\) in pieces. 
    First we begin with \(\diff{}{t}\left\{dh_q\right\}\). 
    Since \(h = (h^1,\ldots,h^k)\) where 
    \(h^i : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R\), we have that 
    \[
        dh_q = \begin{bmatrix}
            dh^1_q \\
            \vdots \\
            dh^k_q
        \end{bmatrix} = \begin{bmatrix}
        \pdiff{h^1}{q_1} & \cdots & \pdiff{h^1}{q_n} \\
        \vdots & \ddots & \vdots \\
        \pdiff{h^k}{q_1} & \cdots & \pdiff{h^k}{q_n}
        \end{bmatrix}
    \]
    The derivative is taken element-wise in each matrix, yielding
    \[
        \diff{}{t}\left\{dh_q\right\} = \begin{bmatrix}
            \sum_{j=1}^n \ppdiff{h^1}{q_1}{q_j}\dot{q}_j & \cdots & \sum_{j=1}^n
            \ppdiff{h^1}{q_n}{q_j}\dot{q}_j \\
            \vdots & \ddots & \vdots \\
            \sum_{j=1}^n \ppdiff{h^k}{q_1}{q_j}\dot{q}_j & \cdots & \sum_{j=1}^n
            \ppdiff{h^k}{q_n}{q_j}\dot{q}_j \\
        \end{bmatrix} + \begin{bmatrix}
            \sum_{l=1}^{n-k}\ppdiff{h^1}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots &
            \sum_{l=1}^{n-k}\ppdiff{h^1}{q_n}{p_{u_l}}\dot{p}_{u_l} \\
            \vdots & \ddots & \vdots \\
            \sum_{l=1}^{n-k}\ppdiff{h^k}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots & 
            \sum_{l=1}^{n-k}\ppdiff{h^k}{q_n}{p_{u_l}}\dot{p}_{u_l} \\
        \end{bmatrix}
    \]
    It is straightforward computation to confirm that each row of this
    derivative can be written in vector form as follows:
    \begin{align*}
        \diff{}{t}\left\{dh^i_q\right\} &= \begin{bmatrix}
        \sum_{j=1}^n \ppdiff{h^i}{q_1}{q_j}\dot{q}_j & \cdots & 
        \sum_{j=1}^n \ppdiff{h^i}{q_n}{q_j}\dot{q}_j 
        \end{bmatrix} + \begin{bmatrix}
        \sum_{l=1}^{n-k}\ppdiff{h^i}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots &
        \sum_{l=1}^{n-k}\ppdiff{h^i}{q_n}{p_{u_l}}\dot{p}_{u_l}
        \end{bmatrix} \\
       &= \dot{q}\tpose \Hess_q\left\{h^i\right\}\tpose + 
       \dot{p}_u\tpose \partial_{p_u}\partial_{q}h^i \\
       &= p\tpose \Minv(q)\Hess_q\left\{h^i\right\}\tpose -
       \left(\frac{1}{2}p\tpose d\Minv_{q_u}(q)(\Id{n-k}\otimes p) + 
       dV_{q_u}(q)\right)\partial_{p_u}\partial_{q}h^i
    \end{align*}
    This means that the entire matrix is given by
    \begin{align*}
        \diff{}{t}\left\{dh_q\right\} &= 
        \left(\Id{n} \otimes 
        \left(p\tpose \Minv(q)\right)\right)
        \Hess_q\left\{h\right\}\tpose -{} \\
      & \left(\frac{1}{2}\Id{n} \otimes 
          \left(p\tpose d\Minv_{q_u}(q)(\Id{n-k}\otimes p)\right)
        + \Id{n}\otimes dV_{q_u}(q)\right)\partial_{p_u}\partial_{q}h
    \end{align*}
    For the next segment, let \(e_j \in \R^n\) be the vector of all zeros,
    except for a single \(1\) at position \(j\). We define
    \(C_1 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n\times n}\) by
    \[
        C_1(q,p) := \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\dot{q}_j 
        = \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\left(e_j\tpose\Minv(q)p\right)
    \]
    Moving on, one can use a similar approach to \(\diff{}{t}\left\{dh_q\right\}\)
    to find the derivative of \(dh_{p_u}\) in matrix form. It is given by
    \begin{align*}
        \diff{}{t}\left\{dh_{p_u}\right\} &= 
        \left( \Id{n} \otimes \left(p\tpose \Minv(q)\right)\right) 
        \partial_q\partial_{p_u}h -{} \\
      & \left(\frac{1}{2}\Id{n} \otimes \left(\pudmat\right) + 
      \Id{n} \otimes \nabla_{q_u}V(q)\right)\Hess_{p_u}\left\{h\right\}\tpose
    \end{align*}
    Now observe that the \(i^\text{th}\) row of
    \[
        \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\}
    \]
    is given by
    \begin{align*}
        \frac{1}{2} \diff{}{t}\left\{p\tpose \pdiff{\Minv}{q_{u_i}}(q)p\right\}
        = p\tpose \pdiff{\Minv}{q_{u_i}}(q) \dot{p} + 
        \frac{1}{2} p\tpose \left( \sum_{j=1}^n \ppdiff{\Minv}{q_{u_i}}{q_j}
        \dot{q}_j \right) p
    \end{align*}
    Defining \(C_2 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n(n-k)\times n}\)
    allows us to write the entire derivative in matrix form:
    \begin{align*}
        & \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\} ={} \\
        & (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q)
        \left(-\frac{1}{2}\pudmat - \nabla_{q_u}V(q) + \simpleB \tau \right) +{}
        \\
        & \frac{1}{2}\left(\Id{n-k}\otimes p\tpose\right) C_2(q,p) p
    \end{align*}
    Finally, we compute the derivative of \(\nabla_{q_u}V(q)\)
    \[
        \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
        \begin{bmatrix}
            \sum_{j=1}^n \ppdiff{V}{q_{u_1}}{q_j}\dot{q}_j \\
            \vdots \\
            \sum_{j=1}^n \ppdiff{V}{q_{u_{n-k}}}{q_j}\dot{q}_j
        \end{bmatrix = \partial_{q_u} \partial_q V(q) \dot{q}
    \]
    and so we get that
    \[
        \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
        \partial_{q_u} \partial_q V(q) \Minv(q)p
    \]
    Putting this all together, we can find the explicit form for \(\ddot{e}\):
    \begin{align*}
        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q)p + dh_q C_1(q,p)p -{}\\
     & \frac{1}{2}dh_q\Minv(q)(\pdmat) + 
        dh_q\Minv(q)\nabla_qV(q) -{}\\
     & \diff{}{t}\left\{dh_{p_u}\right\}
     \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) +{} \\
     & dh_{p_u}(\Id{n-k}\otimes p\tpose)\nabla_{q_u}\Minv(q)\left(
     \frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) -{}\\
     \frac{1}{2}dh_{p_u}(\Id{n-k}\otimes p\tpose)C_2(q,p)p - 
     dh_{p_u} \partial_{q_u}\partial_q V(q)\Minv(q)p +{} \\
     \left(dh_q \Minv(q) - dh_{p_u}(\Id{n-k} \otimes
     p\tpose)\nabla_{q_u}\Minv(q) \right) \simpleB \tau
    \end{align*}
    The VNHC \(h\) is regular iff \(e\) is of relative degree
    \(\{2,\ldots,2\}\), which is true iff \(\tau\) appears in each element of
    \(e\) on the constraint manifold. 
    Letting \(\ddot{e} = E(q,p) + H(q,p)\tau\), this is equivalent to requiring
    that the matrix \(H\) be full rank.
\end{proof}

Using the expression \(\ddot{e} = E(q,p) + H(q,p)\tau\) from the proof of 
Theorem \ref{thm:vnhc-regularity}, a regular VNHC of order \(k\) can be
stabilized by the output-linearizing phase-feedback controller
(\ref{eqn:vnhc-torque-controller}).
\begin{equation}\label{eqn:vnhc-torque-controller}
    \tau(q,p) = -H\inv(q,p)\left(E(q,p) + k_p e + k_d \dot{e}\right)
\end{equation}
where \(k_p, k_d \in \R_{>0}\) are control parameters which can be tuned on the
resulting linear system \(\ddot{e} = -k_p e - k_d\dot{e}\).

@@

Now that we have found a controller to enforce a regular VNHC of order \(k\), we
would like to solve for the closed-loop dynamics. Naturally, these dynamics
should be specified through \((q_u, p_u)\) since \(q_a\) is a function of these
as specified by \(h(q,p_u) = 0\).
Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). This is because
the \(\dot{p}\) dynamics contains the coupling term 
\((\Id{n} \otimes p\tpose)\nabla_{q_u}M(q)p\).

@@

\begin{thm}\label{thm:zero-dynamics}
    Let \(\mathcal{H}\) be a ----- mechanical system satisfying Assumption
    \ref{assm:H-is-simply-actuated}. Let
    \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint manifold
    \(\Gamma\).
    Then, on \(\Gamma\), the closed-loop dynamics are given by
    \begin{equation}\label{eqn:qpu-dynamics}
        \left.\begin{aligned}
            \dot{q}_u &= M(q)p \\
            \dot{p}_u &= -\nabla_{q_u}V(q) \\
            \end{aligned}\right|_{\begin{array}{c}
                h(q,p_u) = 0 \\ 
                pa = g(q_u,p_u) \\
            \end{array}}
    \end{equation}
    where
    \begin{equation}\label{eqn:g-qpu}
    \begin{aligned}
        g(q,p_u) &:= \\
           &\left.\left(dh_q \Minv(q) \simpleB \right)\inv 
        \left(dh_{p_u} \nabla_{q_u}V(q) - dh_q \Minv(q)
        \begin{bmatrix}
            I_{n-k} \\
            \Zmat{k \times (n-k)} \\
        \end{bmatrix} p_u\right)\right|_{h(q,p_u) = 0}
    \end{aligned}
    \end{equation}
\end{thm}
\begin{proof}
%TODO: Complete the proof for this theorem
\end{proof}

@@

\section{Summary of Results}
%TODO: summarize the assumptions and results
